---
title: jsTree htmlwidget
author: Jonathan Sidi
date: '2017-10-24'
slug: jstree
categories: []
tags: [htmlwigdet, rstats]
cran: jsTree
cran_badge: [status, downloads]
gh_repo: metrumresearchgroup/jsTree
gh_badge: [follow, star, issue]
---

jsTree is a R package that is a standalone htmlwidget for the jsTree JavaScript library. It can be run from the R console directly into the Rstudio Viewer, be used in a Shiny application or be part of an Rmarkdown html output.

## Installation

```{r,eval=FALSE}
#CRAN
install.packages('jsTree')

#Github
devtools::install_github('metrumresearchgroup/jsTree')
```

The motivation for the package came from the [shinyTree](https://github.com/trestletech/shinyTree) by Jeff Allen, which has an early version of the JavaScript library hardcoded into a Shiny application.

The input for the htmlwidget is a character vector and the heirarchy is set by the '/', as to mimic the delimiter of a file path. 

The main purpose of this package is vizualize folder hierarchy, much like in the `Files` tab in RStudio. But, instead of looking in local directories it will be used for remote repositories, such as github, bitbucket and svn repositories. This is implemented in the [vcs](https://github.com/metrumresearchgroup/vcs) package and a future post will show utility that package gives.

To show the basic functionality of jsTree we will use a dataset that contains all superfund sites from the [EPA](https://www.epa.gov/superfund/search-superfund-sites-where-you-live). You can access the tibble [here](https://github.com/yonicd/blog/blob/master/data/superfundsite.Rds). The data

```{r,message=FALSE,warning=FALSE}
library(htmlTable)
library(jsTree)
library(widgetframe)
```

```{r,results='asis'}
# Read in the data
superfundsite <- readRDS('../../jstree_data/superfundsite.Rds')

htmlTable::htmlTable(head(superfundsite),rnames=FALSE)
```

## Basic Usage
```{r ex0}
# collapse columns to be delimited by '/'
myx <- apply(superfundsite,1,function(x) paste(x,collapse = '/'))

# call jsTree
j0 <- jsTree(obj = myx)

# place widget in iframe to use in blog post 
# (not needed in regular usage)
frameWidget(j0)
```

## Initialize tree with checked boxes
```{r ex1}
nodestate1 <- superfundsite$State=='California'

j1 <- jsTree(myx,nodestate=nodestate1)

frameWidget(j1)
```

## Tooltips
```{r ex2}
# Use the zipcode as a tooltip on the county name

myx2 <- apply(superfundsite[,-c(5)],1,paste,collapse='/')

zipcode_label <- superfundsite$`Zip Code`
names(zipcode_label) <- superfundsite$County

j2 <- jsTree(myx2,tooltips = zipcode_label)

frameWidget(j2)
```

## Shiny

jsTree can be used in Shiny applications and supplies observers so the Shiny can react to the tree.

<iframe src="https://tranquil-escarpment-51788.herokuapp.com/" style="width: 800px; height: 500px"></iframe>

```{r,eval=FALSE}

ui <- shiny::fluidPage(
  
  shiny::column(width=4,jsTree::jsTreeOutput('tree')),
  
  shiny::column(width=8,shiny::uiOutput('chosen'))
  
)

server <- function(input, output,session) {
  
  network <- shiny::reactiveValues()
  
  shiny::observeEvent(input$tree_update,{
    
    current_selection <- input$tree_update$.current_tree
    if(!is.null(current_selection)) 
      network$tree <- jsonlite::fromJSON(current_selection)
    
  })
  
  shiny::observeEvent(network$tree,{
    
    output$results <- shiny::renderPrint({
      
      str.out <- ''
      
      if(length(network$tree)>0)
        str.out=network$tree
      
      return(str.out)
    })    
  })
  
  output$chosen <- shiny::renderUI({
    
    list(shiny::h3('Selected Items'),
         
    shiny::verbatimTextOutput(outputId = "results"))
    
  })
  
  output$tree <- jsTree::renderJsTree({
    
    nested_string <- myx
    
    jsTree(nested_string)
    
  })
  
}

shinyApp(ui, server)
```